Nested sampling with two objective functions

### 2015-10-26 Paper/ms.tex
Algorithm

\usepackage{dsfont}
\usepackage[utf8]{inputenc}

@@

% Full title of the paper (Capitalized)
\Title{Nested sampling with two or more objective functions}

@@

If we were only interested in a single canonical distribution, for example
with $\beta_1 = 0.3$ and $\beta_2 = 0.7$, we could estimate its normalising
constant and by running standard Nested Sampling with ``likelihood''
$L(\xx) = \exp\left[0.3L_1(\xx) + 0.7L_2(\xx)\right]$. However, usually we
are interested in a range of values for $\beta_1$ and $\beta_2$, and we
want to know the normalisation $Z(\beta_1, \beta_2)$, called the
partition function. This work describes
progress towards solving this class of problems while maintaining the benefits
of Nested Sampling, such as the ability to cope with first-order phase
transitions.

@@

We seek an algorithm for computing the partition function
$Z(\beta_1, \beta_2)$ for a range of values of the $\beta$s, from a single
run. To be considered a variant of Nested Sampling, we require the method
to satisfy the following requirements:
\begin{enumerate}
\item The algorithm should begin with $N$ points drawn from the prior $\pi(\xx)$.
\item The algorithm should seek to explore regions where the values of
$L_1(\xx)$ and $L_2(\xx)$ are greater than what would be expected from the
prior.
\item The algorithm should try to move through a sequence of probability
distributions proportional to $\pi$, but restricted to smaller and smaller
domains for which we can approximately measure the enclosed prior mass.
\item The algorithm should be invariant to monotonic transformations of
$L_1$ and $L_2$, i.e. it should only depend on rankings of $L_1$ and $L_2$
values and not the values themselves.
\end{enumerate}

@@

The algorithm works as follows.

\begin{enumerate}
\item Generate $N$ particles $\{\xx_1, ..., \xx_N\}$ from $\pi(\xx)$.
\item Calculate the {\it corner count} (Equation~\ref{eqn:corner_count})
of each particle.
\item Choose one of the particles which has corner count equal to the
minimum nonzero value. If all particles have a corner count of zero,
the algorithm terminates. Call the selected particle particle $K$.
\item Choose one of the objective functions, and find $K$'s rank in terms
of that objective function.
\item ...
\end{enumerate}

>>>>

\usepackage{dsfont}
\usepackage[utf8]{inputenc}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}

@@

% Full title of the paper (Capitalized)
\Title{Nested sampling with two objective functions}

@@

If we were only interested in a single canonical distribution, for example
with $\beta_1 = 0.3$ and $\beta_2 = 0.7$, we could estimate its normalising
constant and by running standard Nested Sampling with ``likelihood''
$L(\xx) = \exp\left[0.3L_1(\xx) + 0.7L_2(\xx)\right]$. However, usually we
are interested in a range of values for $\beta_1$ and $\beta_2$, and we
want to know the normalisation $Z(\beta_1, \beta_2)$, called the
partition function. This work describes
progress towards solving this class of problems while maintaining the benefits
of Nested Sampling, such as the ability to cope with first-order phase
transitions. The algorithm is designed for problems with two objective
functions, but should generalise to problems with more than two (though perhaps
not more than a few).

@@

We seek an algorithm for computing the partition function
$Z(\beta_1, \beta_2)$ for a range of values of the $\beta$s, from a single
run. To be considered a variant of Nested Sampling, we require the method
to satisfy the following requirements:
\begin{enumerate}
\item The algorithm should begin with $N$ points drawn from the prior $\pi(\xx)$.
\item The algorithm should seek to explore regions where the values of
$L_1(\xx)$ and $L_2(\xx)$ are greater than what would be expected from the
prior.
\item The algorithm should consider a sequence of probability
distributions proportional to $\pi$, but restricted to smaller and smaller
domains for which we can approximately measure the enclosed prior mass.
\item The algorithm should be invariant to monotonic transformations of
$L_1$ and $L_2$, i.e. it should only depend on rankings of $L_1$ and $L_2$
values and not the values themselves.
\end{enumerate}

@@

Here we describe the proposed algorithm, which took three years of
experimentation to develop. Please contact the first author for a list of
failed ideas, some of which might lead to potential alternatives to this
algorithm.

\begin{algorithm}
\begin{algorithmic}
\State Generate $N$ particles $\{\xx_1, ..., \xx_N\}$ from $\pi(\xx)$.
\While{$abc$}
	\State Calculate the {\it corner count} (Equation~\ref{eqn:corner_count})
of each particle.
	\State Choose one of the particles which has corner count equal to the
minimum nonzero value. If all particles have a corner count of zero,
the algorithm terminates. Call the selected particle particle $K$.
	\State Choose one of the objective functions, and find $K$'s rank in terms
of that objective function.
\EndWhile
\end{algorithmic}
\end{algorithm}
